---
title: "TidyTuesday: Verified Oldest People"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(flexdashboard)
library(tidyverse)
library(janitor)
library(plotly)
library(RSQLite)
library(lubridate)
rm(list=ls())
db <- dbConnect(SQLite(), dbname="../DATA/tidytuesday.sqlite3")
```
```{r}
centenarians <- dbGetQuery(db,"SELECT * FROM centenarians;")
centenarians$birth_date <- as.Date(centenarians$birth_date)
centenarians$death_date <- as.Date(centenarians$death_date)
centenarians <- centenarians %>%
  mutate(birth_decade = floor_date(birth_date, "10 years"))
```
The oldest known living person is **Maria Branyas** of Spain, aged 116 years, 85 days. The oldest known **living man is Juan Vicente Pérez** of Venezuela, aged 114 years, 1 day. The 100 oldest women have, on average, lived several years longer than the 100 oldest men.

## Column {data-width=550}


### Centenarians by Gender and Decade


```{r}
centenarians_gender_decade <- centenarians %>%
  mutate(decade = floor_date(birth_date, "10 years")) %>%
  group_by(gender, decade) %>%
  summarise(decade_count = n()) %>%
  arrange(desc(decade_count))

plot_gender_decade <- plot_ly(data = centenarians_gender_decade, x = ~decade, y = ~decade_count, type = "bar", color = ~gender) %>%
  layout(xaxis = list(title = "Decade"), yaxis = list(title = "Count"))

plot_gender_decade


```

### Place of Death or Residence (>1)


```{r}
centenarians_place <- centenarians %>%
  group_by(place_of_death_or_residence) %>%
  summarise(count = n()) %>%
  filter(count > 1) %>%
  arrange(desc(count))

# Define a vector of colors for each category
color_palette <- c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd", "#8c564b", "#e377c2", "#7f7f7f", "#bcbd22", "#17becf")

plot_place <- plot_ly(data = centenarians_place, x = ~count, y = ~reorder(place_of_death_or_residence, -count),
                      type = "bar", color = ~place_of_death_or_residence, colors = color_palette) %>%
  layout(xaxis = list(title = "Count"), yaxis = list(title = "", showticklabels = FALSE)) %>%
  layout(barmode = "stack", showlegend = TRUE) 

plot_place



```

Column {data-width=350}
-----------------------------------------------------------------------

### Centenarians by Still Alive

```{r}
centenarians_alive <- centenarians %>%
  group_by(still_alive) %>%
  summarise(count = n())

plot_alive <- plot_ly(data = centenarians_alive, x = ~still_alive, y = ~count, type = "bar") %>%
  layout(xaxis = list(title = "Still Alive"), yaxis = list(title = "Count"))

plot_alive

```

### Centenarians by Birth Year

```{r}
centenarians_year <- centenarians %>%
  mutate(year = year(birth_date)) %>%
  group_by(year) %>%
  summarise(yr_count = n()) %>%
  arrange(desc(year))

plot_year <- plot_ly(data = centenarians_year, x = ~year, y = ~yr_count, type = "scatter", mode = "lines") %>%
  layout(xaxis = list(title = "Birth Year"), yaxis = list(title = "Count"))

plot_year

```

